Some notes on humanities majors
========================================================


First things first: loading a codebook and some required functions.

```{r}
codebook = read.table("codebook.tsv",sep="\t",stringsAsFactors=F,comment.char="",quote="")
names(codebook) = c("field","fieldname","code","value")
source("functions.R")
#nrow=10000
library(dplyr)
library(ggplot2)
library(matrixStats) # for weighted.mean
tables = src_mysql("ACS",user='bschmidt','password'='Try my raisin Brahms')
persons = tbl(src=tables,"persons")
broadFields = read.table("MajorFields.tsv",sep="\t",header=T) %>% group_by(FOD1P,Discipline) %>% summarize(count=sum(count))
```


First query: collect Age, weighting, and major. We can use this to validate the set by checking it against year-by-year college major numbers.

``` {r}

raw = persons %>% select(AGEP,FOD1P,PWGTP,PWGTP1) %>% collect() %>% encode("FOD1P")


rounding=2

broadFields = read.table("MajorFields.tsv",sep="\t",header=T) %>% group_by(FOD1P,Discipline) %>% summarize(count=sum(count))

table = raw %>%
  mutate(decade=rounding*floor((2012-AGEP)/rounding)) %>% 
  inner_join(broadFields,by="FOD1P") %>% mutate(field=Discipline)


p = table %>% filterToTopN("Discipline",35) %>% filter(!is.na(Discipline))
ageTotals = p %>% group_by(decade) %>% summarise(collegePop = sum(PWGTP))


newtotals = p %>% inner_join(ageTotals) %>% group_by(Discipline,decade) %>% summarise(count=sum(PWGTP/collegePop))

library(scales)
ggplot(newtotals%>% filter(decade<1990)) + geom_area() + aes(x=decade+20,y=count,fill=Discipline) + facet_wrap(~Discipline,scales="free") + scale_y_continuous(labels=percent) 

totals = totals %>% inner_join(broadFields,by="")

```

Something stupid: how do commutes matter?

``` {r}

commutes = persons %>% select(JWMNP,FOD1P,AGEP) %>% collect() %>% encode("FOD1P") %>% filterToTopN("FOD1P",25)
ggplot(commutes) + geom_boxplot(aes(y=JWMNP,x=FOD1P)) + coord_flip() + scale_y_log10()


```


Do overall employment rates differ?
``` {r}
#### Employed at all:

raw = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>22) %>% select(PINCP,AGEP,ESR,FOD1P,SEX,PWGTP)  %>% collect()

employed = raw %>% encode("FOD1P") %>% encode("SEX") %>% encode("ESR")
employed = employed %>% filterToTopN("FOD1P",35)

employStatus = employed %>% filter(AGEP>30,AGEP<39) %>% 
  group_by(FOD1P,SEX) %>% 
  mutate(totals = sum(PWGTP)) %>% 
  group_by("ESR") %>%
  summarize(percent=sum(PWGTP/totals))
  
  
ggplot(employStatus) + 
  geom_bar(aes(fill=ESR,x=FOD1P,y=percent),stat="identity") + coord_flip() + facet_wrap(~SEX)
```

Actual income.
=============

``` {r}

raw = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>22,WKHP>30) %>% 
  #select(PINCP,AGEP,ESR,FOD1P,SEX,PWGTP,PWGTP1,PWGTP10)  %>% 
  collect()


incomes = raw %>% encode("FOD1P") %>% encode("SEX") %>% encode("ESR") %>% mutate(PINCP=as.numeric(PINCP),weight=PWGTP,decade=rounding*floor((2012-AGEP)/rounding),ageGroup = ifelse(AGEP<26,"25 or under",ifelse(AGEP<=33,"26-35",ifelse(AGEP<=45,"36-45",ifelse(AGEP<=60,"46-60","Over 60")))))
#incomes = incomes %>% filterToTopN("FOD1P",35)

incomes = incomes %>% inner_join(broadFields)



#Disciplincomes = incomes %>% filter(AGEP>35,AGEP<=60)%>% group_by(Discipline) %>% summarize(locWeight=sum(weight),income = median(PINCP),meanIncome=weighted.mean(PINCP,weights=weight)) %>% filter(!is.na(Discipline))

summarizingFunction = function(data) {
  value = data %>% 
    group_by(Discipline) %>% 
    summarize(income = median(PINCP),meanIncome=weighted.mean(PINCP,w=weightval)) %>% 
    filter(!is.na(Discipline))
}


errorize = function(data,summarizingFunction) {
  #The summarizing function should include an operation on a weights variable
  
  sampleRun = function(n=1) {
    message("running sample ",n)
    data$weightval = get(paste0("PWGTP",n),data)
    returnval = summarizingFunction(data)
    returnval$run = n 
    silent = returnval
  }
  f = lapply(1:10,sampleRun)
  returnable = f %>% rbind_all()
}

ggplot(returnable) + geom_point(aes(x=meanIncome,y=Discipline),alpha=.33)

ggplot(Disciplincomes) + geom_point(aes(x=income,y=meanIncome)) + coord_equal() + scale_x_continuous(labels=dollar)

ggplot(Disciplincomes) + geom_bar(aes(fill=Discipline,y=income,x=reorder(Discipline,-income)),stat="identity",position="dodge")  + scale_y_continuous("Median income\nSource:American Community Survey",labels=dollar) +coord_flip() + labs(title="Median Incomes, ages 36-60",x="") + theme(axis.text=element_text(size=14),legend.position="None")




```

Income gaps by broad field
---------------------------

``` {r}
incomeTotals = incomes %>% 
    mutate(field=FOD1P) %>% #Uncomment this line to get the results at the individual field level
group_by(field,SEX,ageGroup) %>% summarize(income = median(PINCP),menCounted = sum(SEX=="Male"),womenCounted = sum(SEX=="Female"))

gaps = incomeTotals %>% ungroup() %>% 
  group_by(field,ageGroup,add=F) %>% summarize(gap = (median(income[SEX=="Female"])-median(income[SEX=="Male"]))/median(income[SEX=="Male"]),menCounted=sum(menCounted),womenCounted=sum(womenCounted),income = mean(income)) %>% arrange(-gap)

gapgap = gaps %>% group_by(field) %>% summarize(gapgap = gap[ageGroup=="46-60"]-gap[ageGroup=="25 or under"],menCounted=sum(menCounted),womenCounted=sum(womenCounted)) %>% arrange(-gapgap)

library(scales)
#ggplot(gaps) + geom_bar(aes(x=ageGroup,y=gap,fill=ageGroup),stat="identity",position="dodge")  + scale_y_continuous("Gap between male and female salaries",labels=percent) + facet_wrap(~Discipline,scales="free_x") + labs(title="Gap between median male and female salaries, by college major and age")
```

Per Jan Freeman: do the gaps correspond to the pay?

``` {r}
ggplot(gaps %>% filter(ageGroup=="26-35",menCounted>100,womenCounted>100),aes(x=income,y=gap,label=field,size=menCounted+womenCounted)) + geom_point(alpha=.3) + scale_size_continuous(trans='sqrt',range=c(3,9)) + scale_x_continuous(label=dollar) + geom_smooth(method='lm') + geom_text(data=gaps %>% filter(ageGroup=="26-35",menCounted>1000,womenCounted>1000),size=3) + scale_y_continuous(label=percent) + labs(title="connection between gender gap and income by major: not very strong")
```

Just for 26-35 year olds
-------------------------
``` {r}

younglings = gaps %>% filter(ageGroup=="26-35",!is.na(Discipline)) %>% ungroup() %>% mutate(Discipline = reorder(Discipline,gap))

ggplot(younglings) + geom_bar(aes(x=Discipline,y=-gap,fill=Discipline),stat="identity",position="dodge")  + scale_y_continuous(labels=percent) + labs(title="Gap between median male and female salaries,\nemployed 26-35 year olds, by college major",y="% by which female pay is less than male\nSource: American Community Survey",x="") + coord_flip() + theme(axis.text=element_text(size=14),legend.position="None")


ggplot(incomeTotals %>% filter(ageGroup=="26-35")) + geom_bar(aes(x=FOD1P,y=income,fill=SEX),stat="identity",position="dodge") + coord_flip()

```

The heart of the matter: incomes.

``` {r}
raw = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>20,WKHP>30)  %>% select(PINCP,AGEP,PWGTP,PWGTP1,FOD1P,SEX,SCHL)  %>% collect()
incomes = raw %>% encode("FOD1P") %>% encode("SEX") %>% encode("SCHL")
ggplot(incomes %>% group_by(AGEP,SCHL) %>% 
         summarise(income=mean(PINCP))) + geom_line(aes(x=AGEP,y=income,color=SCHL))

rounding=5

primeYears = incomes %>% filterToTopN("FOD1P",25) %>%
  mutate(decade=rounding*floor((2012-AGEP)/rounding),ageGroup = ifelse(AGEP<26,"25 or under",ifelse(AGEP<=33,"26-35",ifelse(AGEP<=45,"36-45",ifelse(AGEP<=60,"46-60","Over 60"))))) %>% 
  filter(decade >= 1955,decade<=1985)

yearly = primeYears  %>%
  group_by(FOD1P,decade) %>% 
  summarise(income=weighted.mean(as.numeric(PINCP),weights=PWGTP)) %>%
  arrange(-income)

byBins = primeYears %>%
  group_by(FOD1P,ageGroup,SEX) %>%
  summarise(income=weighted.mean(as.numeric(PINCP),weights=PWGTP),median=median(PINCP)) %>%
  group_by(FOD1P,add=F) %>% mutate(overall=median(median)) %>%
  arrange(-overall)

ggplot(byBins) + facet_wrap(~FOD1P) + geom_bar(aes(x=ageGroup,y=median,fill=SEX),position="dodge",stat="identity") + scale_y_continuous(trans="sqrt")

ggplot(byBins) + facet_wrap(~ageGroup) + 
  geom_bar(aes(x=reorder(FOD1P,overall),y=median,fill=SEX),position="dodge",stat="identity") + coord_flip()

persons
```

``` {r}
PHDS = persons %>% filter(SCHL==24,AGEP < 60,AGEP>20,WKHP>30) %>% select(PINCP,AGEP,PWGTP,PWGTP1,FOD2P,SEX) %>% collect()
docs = PHDS %>% encode("FOD2P") %>% encode("SEX") %>% filterToTopN("FOD2P",25) %>%
  mutate(decade=rounding*floor((2012-AGEP)/rounding),ageGroup = ifelse(AGEP<26,"25 or under",ifelse(AGEP<=33,"26-35",ifelse(AGEP<=45,"36-45",ifelse(AGEP<=60,"46-60","Over 60")))))

docs %>% group_by(SEX,FOD2P) %>% 
  summarise(income=weighted.mean(as.numeric(PINCP),weights=PWGTP),num=n(),median = median(PINCP)) %>% 
  arrange(-median)

ggplot(byBins) + facet_wrap(~FOD2P) + geom_line(aes(x=2012-decade,y=income))
```



What industries do historians end up in?

``` {r}
Industries = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>20,WKHP>30) %>% select(SERIALNO,PINCP,AGEP,PWGTP,FOD1P,SEX,INDP) %>% collect()


industries = Industries %>% encode("INDP") %>% encode("FOD1P") %>% mutate(weight=PWGTP)
broadFields = read.table("MajorFields.tsv",sep="\t",header=T) %>% group_by(FOD1P,Discipline) %>% summarize(count=sum(count))
industries = industries %>% inner_join(broadFields)


expectations = industries %>% ungroup() %>% 
  mutate(total=sum(weight),field=FOD1P) %>% 
  group_by(field) %>% 
  mutate(majorCount=sum(weight),majorShare=majorCount/total) %>% 
  group_by(INDP,add=F) %>% 
  mutate(industryCount=sum(weight),industryShare=industryCount/total) %>%
  ungroup() %>% 
  group_by(field,INDP) %>% summarize(count = sum(PWGTP),expected=total[1]*industryShare[1]*majorShare[1],diff=count/expected)


expectations %>% arrange(-count) %>% filter(field=="ENGLISH LANGUAGE AND LITERATURE")

tops = expectations %>% filterToTopN("field",countVariable=quote(count)) %>% filterToTopN("INDP",countVariable=quote(count),n=45)
write.table(tops,file="/tmp/tmp.tsv",sep="\t",quote=F,row.names=F)
ggplot(tops) + geom_tile(aes(x=field,y=INDP,fill=diff)) + scale_fill_gradient2(trans='log') + theme(axis.text.x=element_text(angle=45,hjust=1))


fieldname="Humanities"
ggplot(expectations %>% filter(field==fieldname) %>% arrange(-count) %>% head(50)) + geom_bar(aes(y=count,x=reorder(INDP,count),fill=diff),stat="identity") + theme(axis.text.y=element_text(size=16)) +   scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + coord_flip() + facet_wrap(~field) + geom_point(aes(y=expected,x=reorder(INDP,count)),shape=17) + labs(title=paste("Employment industry of ",tolower(field),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(field),"and red shows an aversion)"),y="Est. number of Americans in the major working in the industry\n(Source: American Community Survey)","Industry of employment") 


```


What jobs do historians end up in?

``` {r}
Industries = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>20,WKHP>30) %>% select(SERIALNO,PINCP,AGEP,PWGTP,FOD1P,SEX,OCCP10,OCCP12) %>% collect()

industries = Industries %>% encode("OCCP10") %>% encode("OCCP12") %>% encode("FOD1P") %>% mutate(occupation=factor(ifelse(is.na(OCCP10),as.character(OCCP12),as.character(OCCP10))))


broadFields = read.table("MajorFields.tsv",sep="\t",header=T) %>% 
  group_by(FOD1P,Discipline) %>% summarize(count=sum(count))
industries = industries %>% inner_join(broadFields)

withWeights = industries %>% inner_join(weights)

expectations = industries %>% mutate(weight=PWGTP) %>% 
  mutate(total=sum(weight),field=FOD1P) %>% 
  group_by(field) %>% 
  mutate(majorCount=sum(weight),majorShare=majorCount/total) %>% 
  group_by(occupation) %>% 
  mutate(industryCount=sum(weight),industryShare=industryCount/total) %>%
  group_by(field,occupation) %>% 
  summarize(count = sum(weight),expected=total[1]*industryShare[1]*majorShare[1],diff=count/expected)



expectations %>% arrange(-count) %>% filter(field=="ENGLISH LANGUAGE AND LITERATURE")

tops = expectations %>% filterToTopN("field",countVariable=quote(count),n=50) %>% filterToTopN("occupation",countVariable=quote(count),n=55)
write.table(tops,file="/tmp/tmp.tsv",sep="\t",quote=F,row.names=F)
ggplot(tops) + geom_tile(aes(x=field,y=occupation,fill=diff)) + scale_fill_gradient2(trans='log') + theme(axis.text.x=element_text(angle=45,hjust=1))
tops
sum(tops$count)/sum(expectations$count)

fieldname="HISTORY"
fieldname="ENGLISH LANGUAGE AND LITERATURE"
ggplot(expectations %>% filter(field==fieldname) %>% arrange(-count) %>% head(50)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") 

ggplot(tops %>% arrange(-count)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field,nrow=2) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") + scale_y_log10()

```
How does ancestry play into this?

``` {r}

ancestries = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>20,WKHP>30) %>% select(SERIALNO,PINCP,AGEP,PWGTP,FOD1P,SEX,ANC1P05,ANC1P12) %>% collect()

industries = ancestries %>% encode("ANC1P12") %>% encode("ANC1P05") %>% encode("FOD1P") %>% mutate(occupation=factor(ifelse(is.na(ANC1P12),as.character(ANC1P05),as.character(ANC1P12))))





broadFields = read.table("MajorFields.tsv",sep="\t",header=T) %>% 
  group_by(FOD1P,Discipline) %>% summarize(count=sum(count))
industries = industries %>% inner_join(broadFields)

withWeights = industries %>% inner_join(weights)

expectations = industries %>% mutate(weight=PWGTP) %>% 
  mutate(total=sum(weight),field=FOD1P) %>% 
  group_by(field) %>% 
  mutate(majorCount=sum(weight),majorShare=majorCount/total) %>% 
  group_by(occupation) %>% 
  mutate(industryCount=sum(weight),industryShare=industryCount/total) %>%
  group_by(field,occupation) %>% 
  summarize(count = sum(weight),expected=total[1]*industryShare[1]*majorShare[1],diff=count/expected)

summarizeWithWeights = function(frame,...) {
  #Should already be somewhat put together.
  dots = list(...)
  dots = lapply(dots,function(defaultFunction) {
    function(weight=quote(PWTGP)) {
      frame
    }
  })
  frame = frame %>% 
    group_by(sample) %>% #Make sure it's grouped by sample
    summarise(count=dots[['count']]())  
#  dots[[".data"]] = frame
  #do.call(summarize,dots)
 # dots
}


head(withWeights) %>% summarizeWithWeights(count=quote(sum(weight)))

expectations %>% arrange(-count) %>% filter(field=="ENGLISH LANGUAGE AND LITERATURE")

tops = expectations %>% filterToTopN("field",countVariable=quote(count),n=50) %>% filterToTopN("occupation",countVariable=quote(count),n=55)
write.table(tops,file="/tmp/tmp.tsv",sep="\t",quote=F,row.names=F)
ggplot(tops) + geom_tile(aes(x=field,y=occupation,fill=diff)) + scale_fill_gradient2(trans='log') + theme(axis.text.x=element_text(angle=45,hjust=1))


fieldname="HISTORY"
fieldname="ENGLISH LANGUAGE AND LITERATURE"
ggplot(expectations %>% filter(field==fieldname) %>% arrange(-count) %>% head(50)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") 

ggplot(tops %>% arrange(-count)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field,nrow=2) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") + scale_y_log10()

```

This amazing map…

``` {r}

ancestries = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>20,WKHP>30) %>% select(SERIALNO,PINCP,AGEP,PWGTP,FOD1P,SEX,ST) %>% collect()

industries = ancestries %>% encode("ST") %>% encode("FOD1P") %>% mutate(occupation=ST)

broadFields = read.table("MajorFields.tsv",sep="\t",header=T) %>% 
  group_by(FOD1P,Discipline) %>% summarize(count=sum(count))
industries = industries %>% inner_join(broadFields)

#withWeights = industries %>% inner_join(weights)

expectations = industries %>% mutate(weight=PWGTP) %>% 
  mutate(total=sum(weight),field=FOD1P) %>% 
  group_by(field) %>% 
  mutate(majorCount=sum(weight),majorShare=majorCount/total) %>% 
  group_by(occupation) %>% 
  mutate(industryCount=sum(weight),industryShare=industryCount/total) %>%
  group_by(field,occupation) %>% 
  summarize(count = sum(weight),expected=total[1]*industryShare[1]*majorShare[1],diff=count/expected)


expectations %>% arrange(-count) %>% filter(field=="ENGLISH LANGUAGE AND LITERATURE")

tops = expectations %>% filterToTopN("field",countVariable=quote(count),n=50) %>% filterToTopN("occupation",countVariable=quote(count),n=55)
write.table(tops,file="/tmp/tmp.tsv",sep="\t",quote=F,row.names=F)
ggplot(tops) + geom_tile(aes(x=field,y=occupation,fill=diff)) + scale_fill_gradient2(trans='log') + theme(axis.text.x=element_text(angle=45,hjust=1))

tops %>% arrange(-diff) %>% group_by(occupation) %>% summarize(topfield=field[1])

fieldname="HISTORY"
fieldname="ENGLISH LANGUAGE AND LITERATURE"
ggplot(expectations %>% filter(field==fieldname) %>% arrange(-count) %>% head(50)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") 

ggplot(tops %>% arrange(-count)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field,nrow=2) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") + scale_y_log10()

```

Will it get you married?


``` {r}

ancestries = persons %>% filter(SCHL>=21,AGEP < 60,AGEP>20,WKHP>30) %>% select(SERIALNO,AGEP,PWGTP,FOD1P,SEX,MAR) %>% collect()

industries = ancestries %>% encode("MAR") %>% encode("FOD1P") %>% mutate(occupation=MAR)

broadFields = read.table("MajorFields.tsv",sep="\t",header=T) %>% 
  group_by(FOD1P,Discipline) %>% summarize(count=sum(count))

industries = industries %>% inner_join(broadFields)

#withWeights = industries %>% inner_join(weights)

head(industries)
summary(lm(occupation=="Divorced"~factor(AGEP)+Discipline,industries,weights=PWGTP))

expectations = industries %>% mutate(weight=PWGTP) %>% 
  mutate(total=sum(weight),field=Discipline) %>% 
  group_by(field) %>% 
  mutate(majorCount=sum(weight),majorShare=majorCount/total) %>% 
  group_by(occupation) %>% 
  mutate(industryCount=sum(weight),industryShare=industryCount/total) %>%
  group_by(field,occupation) %>% 
  summarize(count = sum(weight),expected=total[1]*industryShare[1]*majorShare[1],diff=count/expected)



tops = expectations %>% filterToTopN("field",countVariable=quote(count),n=50) %>% filterToTopN("occupation",countVariable=quote(count),n=55)
ggplot(tops) + geom_tile(aes(x=field,y=occupation,fill=diff)) + scale_fill_gradient2(trans='log') + theme(axis.text.x=element_text(angle=45,hjust=1))




tops %>% arrange(-diff) %>% group_by(occupation) %>% summarize(topfield=field[1])

fieldname="HISTORY"
fieldname="ENGLISH LANGUAGE AND LITERATURE"
ggplot(expectations %>% filter(field==fieldname) %>% arrange(-count) %>% head(50)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") 

ggplot(tops %>% arrange(-count)) + 
  geom_bar(aes(y=count,x=reorder(occupation,count),fill=diff),stat="identity") + 
  theme(axis.text.y=element_text(size=16)) +   
  scale_fill_gradient2("Observed/expected\nratio (blue\nmeans more than expected\n,red less)",trans='log',breaks=as.vector(outer(c(1,2,3,5),c(.01,.1,1,10)))) + 
  coord_flip() + 
  facet_wrap(~field,nrow=2) + 
  geom_point(aes(y=expected,x=reorder(occupation,count)),shape=17) + 
  labs(title=paste("Jobs of ",tolower(fieldname),"majors\n(Expected numbers based on population shown by triangles,\n and color showing ratio, so blue fields show a strong affinity to",tolower(fieldname),"and red shows an aversion)"),y="Est. number of Americans in the major working at job\n(Source: American Community Survey)","Industry of employment") + scale_y_log10()
```
